---
title: "FPT Version Two"
subtitle: "Search Algorithm"
date: last-modified
toc: true
format: 
  html:
    code-fold: true
    grid: 
      margin-width: 350px
      gutter-width: 1em
author: Jessica Helmer
knitr:
    opts_chunk: 
      dev: "ragg_png"
---

::: panel-tabset
#### Setup

```{r}
#| message: false
#| code-fold: show
library(tidyverse)
```

#### Reading

```{r}
#| code-fold: show
dn_dat_c <- readRDS(here::here("Data", "Denominator Neglect", "dn_dat_c.rds"))

bu_dat_h <- readRDS(here::here("Data", "Bayesian Update", "bu_dat_h.rds"))

dn_top_items <- readRDS(here::here("Data", "Denominator Neglect", "top_items.rds"))
```

#### Functions

```{r}
#| code-fold: show

```

#### Cleaning

::: panel-tabset
#### Denominator Neglect

```{r}
dn_dat_c <- dn_dat_c |>
  left_join(dn_top_items |> 
              pivot_longer(c(conf_item, harm_item),
                           names_to = "choice_type", values_to = "item") |>
              select(item, rank),
            by = "item") |>
    mutate(task = "dn.c",
           item = paste0(task, "_", rank),
           # still need time per item
           time = 3.43 / (length(unique(item)) * 2),
           score = correct,
           .keep = "unused") |>
    select(subject_id, sscore, task, item, score, time)
```
:::

#### Bayesian Update

```{r}
bu_dat_h <- bu_dat_h |>
    mutate(task = "bu.h",
           item = paste0(task, "_", unique_trial),
           time = 6.56 / length(unique(item)),
           .keep = "unused") |>
    select(subject_id, sscore, task, item, score, time)
```


#### Combined

Deciding if I want to keep this with the new strategy

```{r}
#| code-fold: show
v2_dat <- rbind(dn_dat_c, 
                bu_dat_h)
```
:::


## Search Algorithm: FPT Version 2

The FPT Version 2 includes two cognitive tasks.

* Denominator Neglect: Combined

* Bayesian Update: Hard

### Methods

This algorithm seeks to maximize the $R^2$ of predicting s-scores per minute for a set of cognitive task items of average testing time `t`.

* For all items of the FPT Version 2 cognitive tasks, calculate the $R^2$ of predicting s-scores of each single item.

* Divide that $R^2$ by the average time it takes to complete an item of that task.

* Identify the item with the maximum $R^2$ per testing time.

* Add that item to the test.

  + Include information of the the task, item number, its $R^2$, and its average testing time.

* For all remaining items, calculate the $R^2$ of a test containing that item and the already included item with s-scores. 

  + Within this test, average the score within task and then within person to get a person score.

* Divide that $R^2$ by the average time it would take to complete the total proposed test.

* Identify the proposed test with the maximum $R^2$ per testing time.

* Continue until the testing time reaches the desired `t`.

Notes:

* For Denominator Neglect, items need to be considered as conflict-harmony pairs to ensure an equal number of conflict and harmony items land on the test.

To manage individual task needs, it may be helpful to identify the maximum correlation per testing time within each task and then compare across tasks. I don't think that will add to any runtime because it will be the same total number of comparisons. 

This means for Denominator Neglect, we can either calculate the correlations across all pairs of possible conflict-harmony items, or we could refer to the top Denominator Neglect items list per the IRT results, and take those as the pairs. To save a little runtime, I'm going to use the pairs list to set the pairs at least to start. 

```{r}
#| eval: false
# this is copied from Cleaning until i find a more elegant solution
dn_dat_c <- dn_dat_c |>
  left_join(dn_top_items |> 
              pivot_longer(c(conf_item, harm_item),
                           names_to = "choice_type", values_to = "item") |>
              select(item, rank),
            by = "item") |>
    mutate(task = "dn.c",
           item = paste0(task, "_", rank),
           time = 3.43 / length(unique(item)),
           score = correct,
           .keep = "unused") |>
    select(subject_id, sscore, task, item, score, time)
```

Task: initialize test dataframe.

```{r}
test <- data.frame(task = as.character(),
                   item = as.character(),
                   test_r2 = as.numeric(),
                   time = as.numeric(),
                   r2_rate = as.numeric())
```

Task: for Denominator Neglect, identify the $R^2$ predicting s-scores for each item pair. And their rates per minute.

```{r}
dn_r2s <- dn_dat_c |>
  summarize(.by = c(subject_id, item),
            score = mean(score),
            sscore = first(sscore),
            time = sum(time),
            task = first(task)) |>
  drop_na() |>
  summarize(.by = item,
            task = first(task),
            test_r2 = summary(lm(sscore ~ score, data = data.frame(sscore, score)))$r.squared,
            time = first(time),
            r2_rate = test_r2 / first(time)) |>
  arrange(desc(r2_rate))
dn_r2s
```

Task: identify the item with the highest $R^2$ rate.

```{r}
dn_max_r2 <- dn_r2s |>
  filter(r2_rate == max(r2_rate))
```

Task: for Bayesian Update, identify the $R^2$ predicting s-scores for each item. And their rates per minute.

```{r}
bu_r2s <- bu_dat_h |>
  
  # this summarize() doesn't actually do anything for BU
  summarize(.by = c(subject_id, item),
            score = mean(score),
            sscore = first(sscore),
            time = sum(time),
            task = first(task)) |>
  drop_na() |>
  summarize(.by = item,
            task = first(task),
            test_r2 = summary(lm(sscore ~ score, data = data.frame(sscore, score)))$r.squared,
            time = first(time),
            r2_rate = test_r2 / first(time)) |>
  arrange(desc(r2_rate))
bu_r2s
```

Task: identify the item with the highest $R^2$

```{r}
bu_max_r2 <- bu_r2s |>
  filter(r2_rate == max(r2_rate))
```

Task: of all tasks, identify the item with the highest $R^2$

```{r}
max_r2 <- rbind(dn_max_r2,
      bu_max_r2) |>
  filter(r2_rate == max(r2_rate))
max_r2
```

Task: add that item to the test

```{r}
test <- rbind(test, max_r2)
```

Now comes the harder part. Need to generalize this to working with (1) already having items within test and (2) each linear model including each task separately.

Starting with Bayesian Update because right away will need separate terms in the linear model.

Task: for Bayesian Update, identify the $R^2$ predicting s-scores for each item. And their rates per minute. Include the currently included items. Okay now its looking like a combined dataframe would be helpful. Joining every rep seems ridiculous. Or maybe I should just join in the set test. I'm convinced we should have a combined set. Not sure if we should only have combined, but I'm going to adjust up above.

```{r}
map(bu_dat_h |> pull(item) |> unique(), 
    ~ rbind(bu_dat_h |>
        filter(item == .x),
      v2_dat |>
        filter(item %in% pull(test, item))) |> 
  summarize(.by = c(task, subject_id),
            score = mean(score),
            sscore = first(sscore),
            time = sum(time),
            task = first(task)) |>
  mutate(.by = subject_id,
         time = sum(time)) |>
  pivot_wider(names_from = task, values_from = score) |>
  summarize(task = str_split_i(.x, "_", 1),
            item = .x,
            test_r2 = summary(lm(sscore ~ dn.c + bu.h,
                                 data = data.frame(sscore, dn.c, bu.h)))$r.squared,
            time = first(time),
            r2_rate = test_r2 / first(time))) |>
  list_rbind() |>
  arrange(desc(r2_rate))
```

Okay, now I'd like to iterate over trying both Bayesian Update items AND Denominator Neglect items. I think I need to change the linear model call so it only includes predictors that are in the data. This could probably be an `if` thing, but I could also do some dynamic formulas. Let's work on it.

```{r}
v2_dat |> 
  filter(!(item %in% pull(test, item))) |>
  pull(item) |>
  unique() |> 
  map(~ rbind(v2_dat |>
                filter(item == .x),
              v2_dat |>
                filter(item %in% pull(test, item))) |> 
        summarize(.by = c(task, subject_id),
                  score = mean(score),
                  sscore = first(sscore),
                  time = sum(time),
                  task = first(task)) |>
        mutate(.by = subject_id,
               time = sum(time)) |> 
        mutate(formula = paste("sscore ~", task |> unique() |> paste(collapse = "+"))) |>
        pivot_wider(names_from = task, values_from = score) |>
        summarize(task = str_split_i(.x, "_", 1),
                  item = .x,
                  test_r2 = summary(lm(as.formula(first(formula)),
                                       data = cur_data()))$r.squared,
                  time = first(time),
                  r2_rate = test_r2 / first(time))) |>
  list_rbind() |>
  arrange(desc(r2_rate))

```

Great. We are able to get out the $R^2$ rate for all items after one item is in the test. And, this works if no items are in the test too.

```{r}
test <- data.frame(task = as.character(),
                   item = as.character(),
                   test_r2 = as.numeric(),
                   time = as.numeric(),
                   r2_rate = as.numeric())

v2_dat |> 
  filter(!(item %in% pull(test, item))) |>
  pull(item) |>
  unique() |> 
  map(~ rbind(v2_dat |>
                filter(item == .x),
              v2_dat |>
                filter(item %in% pull(test, item))) |> 
        summarize(.by = c(task, subject_id),
                  score = mean(score),
                  sscore = first(sscore),
                  time = sum(time),
                  task = first(task)) |>
        mutate(.by = subject_id,
               time = sum(time)) |> 
        mutate(formula = paste("sscore ~", task |> unique() |> paste(collapse = "+"))) |>
        pivot_wider(names_from = task, values_from = score) |>
        summarize(task = str_split_i(.x, "_", 1),
                  item = .x,
                  test_r2 = summary(lm(as.formula(first(formula)),
                                       data = cur_data()))$r.squared,
                  time = first(time),
                  r2_rate = test_r2 / first(time))) |>
  list_rbind() |>
  arrange(desc(r2_rate))
```

So now, we can start putting together this phrase that finds the $R^2$ rates and adding them to the test. Instead of getting a dataframe of all the items' $R^2$ rate, I'd like to start getting back the test itself as items get added.

```{r}
#| eval: false
test <- data.frame(i = as.numeric(),
                   task = as.character(),
                   item = as.character(),
                   test_r2 = as.numeric(),
                   time = as.numeric(),
                   r2_rate = as.numeric())

repeat {
  test <- v2_dat |> 
    filter(!(item %in% pull(test, item))) |>
    pull(item) |>
    unique() |> 
    map(~ rbind(v2_dat |>
                  filter(item == .x),
                v2_dat |>
                  filter(item %in% pull(test, item))) |> 
          summarize(.by = c(task, subject_id),
                    score = mean(score),
                    sscore = first(sscore),
                    time = sum(time),
                    task = first(task)) |>
          mutate(.by = subject_id,
                 time = sum(time)) |> 
          mutate(formula = paste("sscore ~", task |> unique() |> paste(collapse = "+"))) |>
          pivot_wider(names_from = task, values_from = score) |>
          summarize(i = nrow(test) + 1,
                    task = str_split_i(.x, "_", 1),
                    item = .x,
                    test_r2 = summary(lm(as.formula(first(formula)),
                                         data = cur_data()))$r.squared,
                    time = first(time),
                    r2_rate = test_r2 / first(time))) |>
    list_rbind() |>
    filter(r2_rate == max(r2_rate)) |>
    rbind(test)
  
  if (test |> pull(time) |> max() > 1) break
}
```

Ok cool. So that can get us the items that would be included in a one-minute test by optimizing for $R^2$ rate per minute in a linear model assessing items across Denominator Neglect and Bayesian Update. Looks like only Denominator Neglect makes the cut right now. 

But really, I don't know how valuable it is to keep it like that as time-based. I know I said I would do that in the setup, but wouldn't it be better to just run it across all items and get the time / $R^2$ curve? Then we could always just look at desired time cutoffs. 

Is this stepwise strategy better than an all-combinations one? Probably not.

Anyway. Let's get this for all items. I'll just have it iterate over integers through the total potential number of items. 

```{r}
#| eval: false
test <- data.frame(i = as.numeric(),
                   task = as.character(),
                   item = as.character(),
                   test_r2 = as.numeric(),
                   time = as.numeric(),
                   r2_rate = as.numeric())

walk(seq(1, v2_dat |> pull(item) |> unique() |> length()), 
     \(i) {test <<- v2_dat |> 
       filter(!(item %in% pull(test, item))) |>
       pull(item) |>
       unique() |> 
       map(~ rbind(v2_dat |>
                     filter(item == .x),
                   v2_dat |>
                     filter(item %in% pull(test, item))) |> 
             summarize(.by = c(task, subject_id),
                       score = mean(score),
                       sscore = first(sscore),
                       time = sum(time),
                       task = first(task)) |>
             mutate(.by = subject_id,
                    time = sum(time)) |> 
             mutate(formula = paste("sscore ~", task |> unique() |> paste(collapse = "+"))) |>
             pivot_wider(names_from = task, values_from = score) |>
             summarize(i = i,
                       task = str_split_i(.x, "_", 1),
                       item = .x,
                       test_r2 = summary(lm(as.formula(first(formula)),
                                            data = cur_data()))$r.squared,
                       time = first(time),
                       r2_rate = test_r2 / first(time))) |>
       list_rbind() |>
       filter(r2_rate == max(r2_rate)) |>
       rbind(test)}, 
     .progress = T)
test
```


<!-- ```{r} -->
<!-- test -->


<!-- x <- "bu.h_6" -->

<!-- rbind(v2_dat |> -->
<!--         filter(item == x), -->
<!--       v2_dat |> -->
<!--         filter(item %in% pull(test, item))) |>  -->
<!--   summarize(.by = c(task, subject_id), -->
<!--             score = mean(score), -->
<!--             sscore = first(sscore), -->
<!--             time = sum(time), -->
<!--             task = first(task)) |> -->
<!--   mutate(.by = subject_id, -->
<!--          time = sum(time)) |>  -->
<!--   mutate(formula = paste("sscore ~", task |> unique() |> paste(collapse = "+"))) |> -->
<!--   pivot_wider(names_from = task, values_from = score) |> -->
<!--   summarize(task = str_split_i(x, "_", 1), -->
<!--             item = x, -->
<!--             test_r2 = summary(lm(as.formula(first(formula)), -->
<!--                                  data = cur_data()))$r.squared, -->
<!--             added_r2 = first(test_r2) - test |> -->
<!--               filter(i == max(i)) |> -->
<!--               pull(test_r2), -->
<!--             added_time = first(time) - test |> -->
<!--               filter(i == max(i)) |> -->
<!--               pull(time), -->
<!--             time = first(time), -->
<!--             r2_rate = first(added_r2 / added_time)) -->
<!-- ``` -->

```{r}
#| eval: false
test <- data.frame(rep = 0,
                   task = "INIT",
                   item = "INIT",
                   test_r2 = 0,
                   time = 0,
                   added_r2 = 0,
                   added_time = 0,
                   r2_rate = 0)

walk(seq(1, v2_dat |> pull(item) |> unique() |> length()), 
     \(i) {test <<- v2_dat |> 
       filter(!(item %in% pull(test, item))) |>
       pull(item) |>
       unique() |> 
       map(~ rbind(v2_dat |>
                     filter(item == .x),
                   v2_dat |>
                     filter(item %in% pull(test, item))) |> 
             summarize(.by = c(task, subject_id),
                       score = mean(score),
                       sscore = first(sscore),
                       time = sum(time),
                       task = first(task)) |>
             mutate(.by = subject_id,
                    time = sum(time)) |> 
             mutate(formula = paste("sscore ~", task |> unique() |> paste(collapse = "+"))) |>
             pivot_wider(names_from = task, values_from = score) |>
             summarize(rep = i,
                       task = str_split_i(.x, "_", 1),
                       item = .x,
                       test_r2 = summary(lm(as.formula(first(formula)),
                                            data = cur_data()))$r.squared,
                       added_r2 = first(test_r2) - test |>
                         head(1) |>
                         pull(test_r2),
                       added_time = first(time) - test |>
                         head(1) |>
                         pull(time),
                       time = first(time),
                       r2_rate = first(added_r2 / added_time))) |>
       list_rbind() |>
       filter(r2_rate == max(r2_rate)) |>
       rbind(test) |>
       filter(item != "INIT")}, 
     .progress = T)
test

saveRDS(test, here::here("Data", "Search Algorithm", "v2test_dat.rds"))
```


```{r}
test <- readRDS(here::here("Data", "Search Algorithm", "v2test_dat.rds"))
```


```{r}
test |>
  ggplot(aes(x = time, y = test_r2)) +
  geom_line() +
  geom_point(aes(color = task), size = 2) +
  theme_classic() +
  guides(x = guide_axis(cap = "both"),
         y = guide_axis(cap = "both"))


```

Idk why:

```{r}
v2_dat |>
  summarize(.by = c(subject_id, task),
            score = mean(score))
```




Ok something's at least a little off. I'm thinking now that the $R^2$ rate shouldn't be the resulting $R^2$ divided by the new total time but instead the $R^2$ gain divided by the gain in time? Are these the exact same thing?

It's tweaking me out that the $R^2$ decreases throughout. Is that saying that maybe some items are so noisy that they're making the task a worse predictor? Is this a sum score issue?

```{r}

data.frame(y = rnorm(100)) |>
  mutate(x1 = y + rnorm(100, 0),
         x2 = y + rnorm(100),
         x3 = y + rnorm(100),
         x4 = y + rnorm(100),
         x = x1 + x2 + x3 + x4) |>
  lm(y ~ x, data = _) |>
  summary()

data.frame(y = rnorm(100)) |>
  mutate(x1 = y + rnorm(100, 0, 10),
         x2 = y + rnorm(100),
         x3 = y + rnorm(100),
         x4 = y + rnorm(100),
         x = x1 + x2 + x3 + x4) |>
  lm(y ~ x, data = _) |>
  summary()

v2_dat |>
  summarize(.by = c(subject_id, task),
            sscore = first(sscore),
            score = mean(score)) |>
  pivot_wider(names_from = task, values_from = score) |>
  lm(sscore ~ dn.c + bu.h, data = _) |>
  summary() |>
  pluck("r.squared")

```


```{r}
#| eval: false
# todo: conflict-harmony pairs
ranked_items <- data.frame(tank = as.numeric(),
                           task = as.character(),
                           item = as.character(),
                           test_cor = as.numeric(),
                           total_time = as.numeric())

repeat {
  pmap(list(it = v2_dat |> pull(item) |> unique()), \(it) 
       {
         v2_dat |>
           filter(item %in% it | item %in% pull(ranked_items, item)) |>
           summarize(.by = c(task, subject_id),
                     task_score = mean(score),
                     sscore = first(sscore),
                     time = first(time)) |>
           drop_na() |>
           # to-do: change to lm()
           summarize(it = it,
                     test_cor = cor(task_score, sscore),
                     time = first(time))
  }) |> 
    list_rbind() |>
    mutate(cor_rate = test_cor / 
             sum(ranked_items |> pull(total_time), time, na.rm = T)) |>
    filter(cor_rate == max(cor_rate)) |>
    mutate(rank = nrow(ranked_items) + 1,
           task = str_split_i(it , "_", i = 1),
           item = it,
           .keep = "unused",
           .before = everything()) |>
    mutate(total_time = sum(ranked_items |> 
                              arrange(total_time) |>
                              pull(total_time) |>
                              last(), time, na.rm = T)) |>
    rbind(ranked_items) -> ranked_items
  
  if ((ranked_items |> pull(time) |> sum()) > 1) break
}

ranked_items

# dn_top_items |>
#   filter(conf_item == 8 | harm_item == 8) |>
#   select(conf_item, harm_item)

```

## Archive

```{r}
#| eval: false
ranked_items <- data.frame(task = as.character(),
                           item = as.character(),
                           item_cor = as.numeric(),
                           test_cor = as.numeric())


test_cor <- v2_dat |>
  filter(item %in% pull(ranked_items, item)) |>
  summarize(.by = c(task, subject_id),
            task_score = mean(score),
            sscore = first(sscore)) |>
  select(task_score, sscore) |>
  cor()

highest_cor <- c("start" = 0)

v2_dat |>
  filter(item == "bu_h1" | item %in% pull(ranked_items, item)) |>
  summarize(.by = c(task, subject_id),
            task_score = mean(score),
            sscore = first(sscore)) |>
  select(task_score, sscore) |>
  drop_na() |>
  summarize(cor = cor(task_score, sscore)) |>
  pull(cor)

if (.3 > highest_cor) highest_cor <- c("bu_h0" = .2)



```

