---
title: "FPT Version Five"
subtitle: "Search Algorithm"
date: last-modified
toc: true
format: 
  html:
    code-fold: true
    grid: 
      margin-width: 350px
      gutter-width: 1em
author: Jessica Helmer
knitr:
    opts_chunk: 
      dev: "ragg_png"
---

::: panel-tabset
#### Setup

```{r}
#| message: false
#| code-fold: show
library(tidyverse)
```

#### Reading

```{r}
#| code-fold: show
dn_dat_c <- readRDS(here::here("Data", "Denominator Neglect", "dn_dat_c.rds"))

bu_dat_h <- readRDS(here::here("Data", "Bayesian Update", "bu_dat_h.rds"))

admc.dr_dat <- readRDS(here::here("Data", "ADMC Decision Rules", "admc.dr_dat.rds"))

ns_dat <- readRDS(here::here("Data", "Number Series", "ns_dat.rds"))

dn_dat_s <- readRDS(here::here("Data", "Denominator Neglect", "dn_dat_s.rds"))

bu_dat_e <- readRDS(here::here("Data", "Bayesian Update", "bu_dat_e.rds"))

dn_top_items_c <- readRDS(here::here("Data", "Denominator Neglect", "top_items.rds"))
dn_top_items_s <- readRDS(here::here("Data", "Denominator Neglect", "top_items_s.rds"))
```

#### Functions

```{r}
#| code-fold: show

```

#### Cleaning

::: panel-tabset

#### Denominator Neglect

```{r}
dn_dat_c <- dn_dat_c |>
  left_join(dn_top_items_c |> 
              pivot_longer(c(conf_item, harm_item),
                           names_to = "choice_type", values_to = "item") |>
              select(item, rank),
            by = "item") |>
    mutate(task = "dn.c",
           item = paste0(task, "_", rank),
           # still need time per item
           time = 3.43 / (length(unique(item)) * 2),
           score = correct,
           .keep = "unused") |>
    select(subject_id, sscore, task, item, score, time)

dn_dat_s <- dn_dat_s |>
  left_join(dn_top_items_s |> 
              pivot_longer(c(conf_item, harm_item),
                           names_to = "choice_type", values_to = "item") |>
              select(item, rank),
            by = "item") |>
    mutate(task = "dn.s",
           item = paste0(task, "_", rank),
           # still need time per item
           time = 4.45 / (length(unique(item)) * 2),
           score = correct,
           .keep = "unused") |>
    select(subject_id, sscore, task, item, score, time)
```

#### Bayesian Update

```{r}
bu_dat_h <- bu_dat_h |>
    mutate(task = "bu.h",
           item = paste0(task, "_", unique_trial),
           time = 6.56 / length(unique(item)),
           .keep = "unused") |>
    select(subject_id, sscore, task, item, score, time)

bu_dat_e <- bu_dat_e |>
    mutate(task = "bu.e",
           item = paste0(task, "_", unique_trial),
           time = 7.26 / length(unique(item)),
           .keep = "unused") |>
    select(subject_id, sscore, task, item, score, time)
```

#### ADMC: Decision Rules

```{r}
admc.dr_dat <- admc.dr_dat |>
    mutate(task = "admc.dr",
           item = paste0(task, "_", admc_id),
           time = 5.87 / length(unique(item)),
           .keep = "unused") |>
    select(subject_id, sscore, task, item, score, time)
```

#### Number Series

```{r}
ns_dat <- ns_dat |>
    mutate(task = "ns",
           item = paste0(task, "_", ns_id),
           time = 3.70 / length(unique(item)),
           .keep = "unused") |>
    select(subject_id, sscore, task, item, score, time)
```

#### Combined

```{r}
#| code-fold: show
v5_dat <- rbind(dn_dat_c, 
                bu_dat_h,
                admc.dr_dat,
                ns_dat,
                dn_dat_s,
                bu_dat_e)
```

Note that all time information is based on the median time (from FPT preprint) for the entire task divided by number of items.
:::

:::

## Search Algorithm: FPT Version 5

The FPT Version 5 includes five cognitive tasks.

* Denominator Neglect: Combined

* Bayesian Update: Hard

* ADMC: Decision Rules

* Number Series

* Cognitive Reflection Test

* Matrix Reasoning

* Denominator Neglect: Separate

* Bayesian Update: Easy

I don't know what to do with the CRT as of now, so this will only have Denominator Neglect: Combined, Bayesian Update: Hard, ADMC: Decision Rules, Number Series.

I also don't know what to do with Raven.

### Methods

This algorithm seeks to maximize the $R^2$ of predicting s-scores per minute for a set of cognitive task items of average testing time `t`.


```{r}
#| eval: false
test <- data.frame(rep = 0,
                   task = "INIT",
                   item = "INIT",
                   test_r2 = 0,
                   time = 0,
                   added_r2 = 0,
                   added_time = 0,
                   r2_rate = 0)

walk(seq(1, v5_dat |> pull(item) |> unique() |> length()), 
     \(i) {test <<- v5_dat |> 
       filter(!(item %in% pull(test, item))) |>
       pull(item) |>
       unique() |> 
       map(~ rbind(v5_dat |>
                     filter(item == .x),
                   v5_dat |>
                     filter(item %in% pull(test, item))) |> 
             summarize(.by = c(task, subject_id),
                       score = mean(score),
                       sscore = first(sscore),
                       time = sum(time),
                       task = first(task)) |>
             mutate(.by = subject_id,
                    time = sum(time)) |> 
             mutate(formula = paste("sscore ~", task |> unique() |> paste(collapse = "+"))) |>
             pivot_wider(names_from = task, values_from = score) |>
             summarize(rep = i,
                       task = str_split_i(.x, "_", 1),
                       item = .x,
                       test_r2 = summary(lm(as.formula(first(formula)),
                                            data = pick(everything())))$r.squared,
                       added_r2 = first(test_r2) - test |>
                         head(1) |>
                         pull(test_r2),
                       added_time = first(time) - test |>
                         head(1) |>
                         pull(time),
                       time = first(time),
                       r2_rate = first(added_r2 / added_time))) |>
       list_rbind() |>
       filter(r2_rate == max(r2_rate)) |>
       rbind(test) |>
       filter(item != "INIT")}, 
     .progress = T)
test

saveRDS(test, here::here("Data", "Search Algorithm", "v5test_dat.rds"))
```

```{r}
test <- readRDS(here::here("Data", "Search Algorithm", "v5test_dat.rds"))
```


```{r}
#| fig-width: 6
#| fig-height: 3.5
test |>
  ggplot(aes(x = time, y = test_r2)) +
  geomtextpath::geom_textsegment(data = filter(test, test_r2 == max(test_r2)),
               aes(x = -Inf, xend = time, y = test_r2, yend = test_r2, label = "PEAK"),
               color = "gray80", linewidth = .3, linetype = "longdash") +
  geom_segment(data = filter(test, test_r2 == max(test_r2)),
               aes(y = -Inf, yend = test_r2, x = time),
               color = "gray80", linewidth = .3, linetype = "longdash") +
  geom_line() +
  geom_point(aes(color = task), size = 2) +
  guides(x = guide_axis(cap = "both"),
         y = guide_axis(cap = "both")) +
  labs(x = "Time (minutes)", y = "Test R^2") +
  theme_classic()
```

So this method still uses a pretty crude way of measuring time. It does not account for (1) the time it takes to read instructions / do practice items, (2) any between-item difference in average time. Could

* Make doing an item from a new task account for its upfront time cost

* Other stuff
