---
title: "Bayesian Update"
date: last-modified
toc: true
format: 
  html:
    code-fold: true
    grid: 
      margin-width: 350px
      gutter-width: 1em
author: Jessica Helmer
knitr:
    opts_chunk: 
      dev: "ragg_png"
---

::: panel-tabset
#### Setup

```{r}
#| message: false
#| code-fold: show
library(tidyverse)
```

#### Reading

```{r}
#| code-fold: show
bu_full <- rio::import("https://raw.githubusercontent.com/forecastingresearch/fpt/refs/heads/main/data_cognitive_tasks/task_datasets/data_bayesian_update.csv")

anchoraig_full <- rio::import("https://raw.githubusercontent.com/forecastingresearch/fpt/refs/heads/main/data_cognitive_tasks/metadata_tables/task_aig_version.csv")

dn_dat <- readRDS(here::here("Data", "Denominator Neglect", "dn_dat.rds"))

dn_topitems <- readRDS(here::here("Data", "Denominator Neglect", "top_items.rds"))
```

#### Functions

```{r}
#| code-fold: show
# calculates the correlation between a randomly sample group of `n_items` and s-scores
bu_func <- function(data, n_items, rep = 1:1) {
  pmap(expand_grid(n_items = n_items, rep = rep), 
          \(n_items, rep) {
            
            data |>
              filter(unique_trial %in% sample(unique(unique_trial), n_items)) |>
              
              # get person-level scores for both `sscore` and BU (`score`)
              summarize(.by = subject_id,
                        sscore = first(sscore),
                        score = mean(score),
                        n_items = n()) |>
              select(sscore, score) |>
              drop_na() |>
              summarize(n_items = n_items,
                        rep = rep,
                        cor = cor(sscore, score) |> abs())
          }) |>
    list_rbind()
}

# calculates the R^2 of a linear model with `n_items` BU items and the top five DN items
bu_r2_func <- function(data, n_items, rep = 1) {
  pmap(expand_grid(n_items = n_items, rep = rep), 
       \(n_items, rep) {
         
         data |>
           filter(unique_trial %in% sample(unique(unique_trial), n_items)) |>
           summarize(.by = subject_id,
                     sscore = first(sscore),
                     correct = first(correct),
                     score = mean(score),
                     n_items = n()) |>
           select(sscore, score, correct) |>
           drop_na() |>
           summarize(n_items = n_items,
                     rep = rep,
                     r2 = lm(sscore ~ score + correct,
                             data = data.frame(sscore, score, correct)) |> 
                       summary() |>
                       pluck("r.squared"))
       }) |>
    list_rbind()
}
```

#### Cleaning

::: panel-tabset
#### Easy
```{r}
#| code-fold: show
anchor_ids_e <- anchoraig_full %>%
  filter(task == "bayesian_update_easy" |
           task == "bayesian_update_hard") %>%
  mutate(task_version = factor(ifelse(task == "bayesian_update_easy", "easy", "hard"))) %>%
  select(!task) %>%
  
  # because participants did more than one version in a session, need to specify both anchor and AIG both here and in the following assignment to get just the specific quadrant
  filter(AIG_version == "anchor" & task_version == "easy") %>%
  pull(session_id)

bu_dat_e <- bu_full %>%
  arrange(session_id, unique_trial) %>%
  
  # only anchor version and only easy version
  filter(session_id %in% anchor_ids_e & version == "easy") |>
  select(session_id, unique_trial, score, ball_split) |>
  left_join(rio::import("https://raw.githubusercontent.com/forecastingresearch/fpt/refs/heads/main/data_cognitive_tasks/metadata_tables/session.csv"),
            by = "session_id") |>
  left_join(rio::import("https://raw.githubusercontent.com/forecastingresearch/fpt/refs/heads/main/data_forecasting/processed_data/scores_quantile.csv") |>
            select(subject_id, sscore_standardized) |>
              
            # mean sscores per person
            summarize(.by = subject_id,
                      sscore = mean(sscore_standardized)),
            by = "subject_id") |>
  select(subject_id, unique_trial, ball_split, score, sscore) |>
  
  # mean score per BU item (over the five trials)
  # suggestion (maybe): do some trial-level stuff at some point?
  summarize(.by = c(subject_id, unique_trial),
            score = mean(score),
            sscore = first(sscore),
            ball_split = first(ball_split))

saveRDS(bu_dat_e, here::here("Data", "Bayesian Update", "bu_dat_e.rds"))
```

#### Hard
```{r}
#| code-fold: show
anchor_ids_h <- anchoraig_full %>%
  filter(task == "bayesian_update_easy" |
           task == "bayesian_update_hard") %>%
  mutate(task_version = factor(ifelse(task == "bayesian_update_easy", "easy", "hard"))) %>%
  select(!task) %>%
  
  # because participants did more than one version in a session, need to specify both anchor and AIG both here and in the following assignment to get just the specific quadrant
  filter(AIG_version == "anchor" & task_version == "hard") %>%
  pull(session_id)

bu_dat_h <- bu_full %>%
  arrange(session_id, unique_trial) %>%
  
  # only anchor version and only easy version
  filter(session_id %in% anchor_ids_h & version == "hard") |>
  select(session_id, unique_trial, score, ball_split) |>
  left_join(rio::import("https://raw.githubusercontent.com/forecastingresearch/fpt/refs/heads/main/data_cognitive_tasks/metadata_tables/session.csv"),
            by = "session_id") |>
  left_join(rio::import("https://raw.githubusercontent.com/forecastingresearch/fpt/refs/heads/main/data_forecasting/processed_data/scores_quantile.csv") |>
            select(subject_id, sscore_standardized) |>
              
            # mean sscores per person
            summarize(.by = subject_id,
                      sscore = mean(sscore_standardized)),
            by = "subject_id") |>
  select(subject_id, unique_trial, ball_split, score, sscore) |>
  
  # mean score per BU item (over the five trials)
  # suggestion (maybe): do some trial-level stuff at some point?
  summarize(.by = c(subject_id, unique_trial),
            score = mean(score),
            sscore = first(sscore),
            ball_split = first(ball_split))

saveRDS(bu_dat_h, here::here("Data", "Bayesian Update", "bu_dat_h.rds"))
```
:::

:::


```{r from script}
library(tidyverse)

bu_full <- rio::import("https://raw.githubusercontent.com/forecastingresearch/fpt/refs/heads/main/data_cognitive_tasks/task_datasets/data_bayesian_update.csv")

anchoraig_full <- rio::import("https://raw.githubusercontent.com/forecastingresearch/fpt/refs/heads/main/data_cognitive_tasks/metadata_tables/task_aig_version.csv")

bu_full |>
  summarize(.by = trial,
            draw = list(unique(current_draw)))

anchor_ids_h <- anchoraig_full %>%
  filter(task == "bayesian_update_hard" & AIG_version == "anchor") |>
  pull(session_id)

bu_full %>%
  filter(session_id %in% anchor_ids_h & version == "hard") |>
  summarize(.by = trial,
            ball_split = unique(ball_split),
            draw = unique(current_draw))

```

The following analysis only focuses on Bayesian Update: Easy

### Correlations Between Bayesian Update and S-Scores

The overall correlation between mean scores on Bayesian Update and s-scores was `r (bu_dat_e |> summarize(.by = subject_id, sscore = first(sscore), score = mean(score)) |> drop_na() |> select(sscore, score) |> cor() |> round(2))[1, 2]`.

```{r}
#| column: margin
#| fig-height: 2.5
#| fig-width: 2.5
#| warning: false
bu_dat_e |>
  filter(unique_trial %in% sample(unique(unique_trial), 8)) |>
  summarize(.by = subject_id,
            sscore = first(sscore),
            score = mean(score)) |>
  ggplot(aes(x = score, y = sscore)) +
  geom_point() +
  # chore: italicize "r"
  geom_label(label = paste("r =", (bu_dat_e |>
                                       summarize(.by = subject_id, sscore = first(sscore), score = mean(score)) |> 
                                       drop_na() |>
                                       select(sscore, score) |>
                                       cor() |> 
                                      round(2))[1, 2]),
             x = Inf, y = Inf, hjust = 1, vjust = 1) +
  labs(y = "S-Score", x = "Bayesian Update") +
  guides(x = guide_axis(cap = "both"), y = guide_axis(cap = "both")) +
  theme_classic() +
  theme(aspect.ratio = 1)
```

We randomly selected $n$ items where $n \in [1, 8]$ and calculated the correlations between person-level mean scores on those $n$ items and s-scores, doing 100 replications per $n$.

```{r}
#| eval: false
bu_cors <- bu_func(bu_dat_e, 1:8, 1:100)
saveRDS(bu_cors, here::here("Data", "Bayesian Update", "bu_cors.rds"))
```


```{r}
bu_cors <- readRDS(here::here("Data", "Bayesian Update", "bu_cors.rds"))

bu_cors |>
    summarize(.by = c(n_items),
            across(cor, list(mean = mean,
                             lower = ~ quantile(.x, 0.05, names = FALSE),
                             upper = ~ quantile(.x, 0.95, names = FALSE)))) |>
  ggplot(aes(x = n_items, y = cor_mean)) +
  geom_line(linewidth = 1.5, color = "slategray") +
  geom_ribbon(aes(ymin = cor_lower, ymax = cor_upper),
              alpha = .3, fill = "slategray") +
  scale_x_continuous(breaks = c(1, 3, 6, 8)) +
  scale_y_continuous(breaks = c(.2, .4, .6)) +
  coord_cartesian(ylim = c(.2, .6)) +
  labs(y = "Correlation with S-Score", x = "Number of Items") +
  guides(x = guide_axis(cap = "both"), y = guide_axis(cap = "both")) +
  theme_classic() 
```

## Correlations Between Bayesian Update and S-Score Residuals

We first residualized out the effect of the top five Denominator Neglect items.

```{r}
dn_topfive <- dn_topitems |>
  pivot_longer(c(conf_item, harm_item),
               names_to = "choice_type", values_to = "item") |>
  filter(rank <= 5) |>
  pull(item)

dn_topfivedat <- dn_dat |> 
  filter(item %in% dn_topfive) |> 
  summarize(.by = subject_id,
            sscore = first(sscore),
            correct = mean(correct)) |> # alert alert
  drop_na()

dn_resids <- data.frame(subject_id = dn_topfivedat$subject_id,
      residuals = lm(sscore ~ correct, data = dn_topfivedat)$residuals)
```

The overall correlation between mean scores on Bayesian Update and s-score residuals was `r (bu_dat_e |> left_join(dn_resids, by = "subject_id") |> mutate(sscore = residuals) |> summarize(.by = subject_id, sscore = first(sscore), score = mean(score)) |> drop_na() |> select(sscore, score) |> cor() |> round(2))[1, 2]`.

```{r}
#| column: margin
#| fig-height: 2.5
#| fig-width: 2.5
#| warning: false
bu_dat_e |>
  left_join(dn_resids, by = "subject_id") |>
  mutate(sscore = residuals) |>
  summarize(.by = subject_id,
            sscore = first(sscore),
            score = mean(score)) |>
  ggplot(aes(x = score, y = sscore)) +
  geom_point() +
  # chore: italicize "r"
  geom_label(label = paste("r =", (bu_dat_e |>
  left_join(dn_resids, by = "subject_id") |>
  mutate(sscore = residuals) |>
                                       summarize(.by = subject_id, sscore = first(sscore), score = mean(score)) |> 
                                       drop_na() |>
                                       select(sscore, score) |>
                                       cor() |> 
                                      round(2))[1, 2]),
             x = Inf, y = Inf, hjust = 1, vjust = 1) +
  labs(y = "S-Score Residuals", x = "Bayesian Update") +
  guides(x = guide_axis(cap = "both"), y = guide_axis(cap = "both")) +
  theme_classic() +
  theme(aspect.ratio = 1)
```

```{r}
#| eval: false
bu_cors_postdn <- bu_dat_e |>
  left_join(dn_resids, 
            by = "subject_id") |>
  mutate(sscore = residuals) |>
  bu_func(1:8, 1:100)

saveRDS(bu_cors_postdn, here::here("Data", "Bayesian Update", "bu_cors_postdn.rds"))
```

```{r}
bu_cors_postdn <- readRDS(here::here("Data", "Bayesian Update", "bu_cors_postdn.rds"))

bu_cors_postdn |>
  summarize(.by = c(n_items),
            across(cor, list(mean = mean,
                             lower = ~ quantile(.x, 0.05, names = FALSE),
                             upper = ~ quantile(.x, 0.95, names = FALSE)))) |>
  ggplot(aes(x = n_items, y = cor_mean)) +
  geom_line(linewidth = 1.5, color = "slategray") +
  geom_ribbon(aes(ymin = cor_lower, ymax = cor_upper),
              alpha = .3, fill = "slategray") +
  scale_x_continuous(breaks = c(1, 3, 6, 8)) +
  scale_y_continuous(breaks = c(.2, .3, .4)) +
  coord_cartesian(ylim = c(.2, .4)) +
  labs(y = "Correlation with S-Score Residuals", x = "Number of Items",
       caption = "Top five Denominator Neglect items\nresidualized out of s-scores") +
  guides(x = guide_axis(cap = "both"), y = guide_axis(cap = "both")) +
  theme_classic() 
```

### $R^2$ Post-DN and BN

To see how successful the combinations of the five Denominator Neglect items and $n$ Bayesian Update items are at capturing forecasting ability, here is the $R^2$ of a simple linear model with increasing $n$ of Bayesian Update items in the person-level means scores.

```{r}
#| eval: false
bu_r2 <- bu_dat_e |>
           left_join(dn_topfivedat,
                     by = c("subject_id", "sscore")) |>
  bu_r2_func(1:8, 1:100)

saveRDS(bu_r2, here::here("Data", "Bayesian Update", "bu_r2.rds"))
```

```{r}
bu_r2 <- readRDS(here::here("Data", "Bayesian Update", "bu_r2.rds"))

bu_r2 |>
  summarize(.by = c(n_items),
            across(r2, list(mean = mean,
                             lower = ~ quantile(.x, 0.05, names = FALSE),
                             upper = ~ quantile(.x, 0.95, names = FALSE)))) |>
  ggplot(aes(x = n_items, y = r2_mean)) +
  geom_line(linewidth = 1.5, color = "slategray") +
  geom_ribbon(aes(ymin = r2_lower, ymax = r2_upper),
              alpha = .3, fill = "slategray") +
  scale_x_continuous(breaks = c(1, 3, 6, 8)) +
  scale_y_continuous(breaks = c(.2, .3, .4)) +
  coord_cartesian(ylim = c(.2, .4)) +
  labs(y = "R^2", x = "Number of BU Items",
       caption = "Top five Denominator Neglect items included in model") +
  guides(x = guide_axis(cap = "both"), y = guide_axis(cap = "both")) +
  theme_classic() 
```














