---
title: "LASSO Regression"
date: last-modified
toc: true
format: 
  html:
    code-fold: true
    grid: 
      margin-width: 350px
      gutter-width: 1em
author: Jessica Helmer
knitr:
    opts_chunk: 
      dev: png
      dev.args:
        bg: transparent
lightbox: auto
---

::: panel-tabset
#### Setup

```{r}
#| message: false
#| code-fold: show
library(tidyverse)
library(tidymodels)
library(vip)
library(gt)

set_theme(theme_classic(base_size = 16, paper = "#eceadf"))
```

#### Reading

```{r}
#| code-fold: show
v5_dat <- readRDS(here::here("Data", "v5_dat.rds"))

startup_costs <- readRDS(here::here("Data", "startup_costs.rds"))
```

#### Functions

```{r}
#| code-fold: show
source(here::here("Scripts", "get_startup.R"))
```

#### Cleaning

```{r}
v5_dat_wide <- v5_dat |>
  # condensing denominator neglect scores down to one score per item pair
  mutate(.by = c(subject_id, item),
         score = ifelse(task == "dn.c" | task == "dn.s",
                        mean(score),
                        score)) |>
  unique() |>
  select(-c(time, task)) |>
  pivot_wider(names_from = "item", values_from = "score",
              id_cols = c(subject_id, sscore)) |>
  # alert alert !!
  drop_na()
```
:::

Splitting data

```{r}
set.seed(12345)

# splitting data
v5_dat_split <- initial_split(v5_dat_wide)

v5_dat_train <- training(v5_dat_split)
v5_dat_test <- testing(v5_dat_split)

# building "recipe" of all items regressed on sscore
v5_dat_recipe <- recipe(sscore ~ ., data = v5_dat_train) |>
  # not including subject_id
  update_role(subject_id, new_role = "ID") |>
  # and z-scoring first
  step_normalize(all_numeric(), -all_outcomes())

wf <- workflow() |>
  add_recipe(v5_dat_recipe)

v5_dat_boot <- bootstraps(v5_dat_train)
```


## No Tuning

::::: {.callout-note collapse="true"}
## No Tuning Models

:::: {.columns}
::: {.column width="50%"}
### LASSO

```{r}
#| message: false
# linear regression with glmnet, mixture = 1 to do LASSO
lasso_spec <- linear_reg(penalty = 0.1, mixture = 1) |>
  set_engine("glmnet")

# fitting model with training data
lasso_fit <- wf |>
  add_model(lasso_spec) |>
  fit(data = v5_dat_train)

# extracting estimates
lasso_estimates <- lasso_fit |>
  extract_fit_parsnip() |>
  tidy()

# display estimates
# lasso_estimates |> 
#   filter_out(when_any(estimate == 0, 
#                       term == "(Intercept)")) |>
#   mutate(task = term |> str_split_i("_", 1)) |>
#   select(task, item = term, estimate)
```

```{r}
#| fig-height: 7.5
lasso_estimates |>
  filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  ggplot(aes(y = fct_reorder(term, abs(estimate)), x = abs(estimate),
             color = task)) +
  geom_point() +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  labs(x = "Absolute Estimate", y = NULL, color = "Task",
       title = "Non-Zero Estimates") +
  theme(aspect.ratio = 3)
```

```{r}
lasso_test_time <- lasso_estimates |> filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  inner_join(v5_dat |> select(item, time) |> unique(),
            join_by("term" == "item")) |>
  summarize(.by = task,
            time = sum(time) + get_startup(first(task))) |>
  summarize(time = sum(time)) |>
  pull(time)
```

Time of untuned LASSO test: `r lasso_test_time |> round(2)` minutes.

:::

::: {.column width="50%"}
### Elastic Net

```{r}
#| message: false

# # splitting data
# v5_dat_split <- initial_split(v5_dat_wide)
# 
# v5_dat_train <- training(v5_dat_split)
# v5_dat_test <- testing(v5_dat_split)
# 
# # building "recipe" of all items regressed on sscore
# v5_dat_recipe <- recipe(sscore ~ ., data = v5_dat_train) |>
#   # not including subject_id
#   update_role(subject_id, new_role = "ID") |>
#   # and z-scoring first
#   step_normalize(all_numeric(), -all_outcomes())

# linear regression with glmnet, mixture = 0.5 to do elastic net
elasticnet_spec <- linear_reg(penalty = 0.1, mixture = 0.5) |>
  set_engine("glmnet")

# wf <- workflow() |>
#   add_recipe(v5_dat_recipe)

# fitting model with training data
lasso_fit <- wf |>
  add_model(elasticnet_spec) |>
  fit(data = v5_dat_train)

# extracting estimates
lasso_estimates <- lasso_fit |>
  extract_fit_parsnip() |>
  tidy()

# display estimates
# lasso_estimates |> 
#   filter_out(when_any(estimate == 0, 
#                       term == "(Intercept)")) |>
#   mutate(task = term |> str_split_i("_", 1)) |>
#   select(task, item = term, estimate)
```

```{r}
#| fig-height: 9
lasso_estimates |>
  filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  ggplot(aes(y = fct_reorder(term, abs(estimate)), x = abs(estimate),
             color = task)) +
  geom_point() +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  labs(x = "Absolute Estimate", y = NULL, color = "Task",
       title = "Non-Zero Estimates") +
  theme(aspect.ratio = 3)
```

```{r}
lasso_test_time <- lasso_estimates |> filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  inner_join(v5_dat |> select(item, time) |> unique(),
            join_by("term" == "item")) |>
  summarize(.by = task,
            time = sum(time) + get_startup(first(task))) |>
  summarize(time = sum(time)) |>
  pull(time)
```

Time of untuned elastic net test: `r lasso_test_time |> round(2)` minutes.

:::
::::

:::::

## Tuned Models

:::: {.columns}
::: {.column width="50%"}
### LASSO Tuning

Tuning the `penalty` argument.

```{r}
# tune penalty instead of fixing it
# mixture = 1 for LASSO
lasso_tune_spec <- linear_reg(penalty = tune(), mixture = 1) |>
  set_engine("glmnet")
```

```{r}
#| eval: true
# set.seed(1234)
# 
# # splitting data
# v5_dat_split <- initial_split(v5_dat_wide)
# 
# v5_dat_train <- training(v5_dat_split)
# v5_dat_test <- testing(v5_dat_split)




lambda_grid <- grid_regular(penalty(), levels = 50)

doParallel::registerDoParallel()

lasso_grid <- tune_grid(
  wf |> add_model(lasso_tune_spec),
  resamples = v5_dat_boot,
  grid = lambda_grid,
  metrics = metric_set(rmse)
)

saveRDS(lasso_grid, here::here("Data", "lasso-estimates_tuning.rds"))
```


<!-- ```{r} -->
<!-- #| column: margin -->


<!-- # lasso_grid |> -->
<!-- #   collect_metrics() |> -->
<!-- #   ggplot(aes(penalty, mean, color = .metric)) + -->
<!-- #   geom_errorbar(aes( -->
<!-- #     ymin = mean - std_err, -->
<!-- #     ymax = mean + std_err), -->
<!-- #   alpha = 0.5) + -->
<!-- #   geom_line(linewidth = 1.5, na.rm = T) + -->
<!-- #   facet_wrap(~.metric, scales = "free", nrow = 2) + -->
<!-- #   scale_x_log10() + -->
<!-- #   theme(legend.position = "none") -->
<!-- ``` -->


```{r}
#| fig-height: 8
lasso_grid <- readRDS(here::here("Data", "lasso-estimates_tuning.rds"))

lowest_rmse <- lasso_grid |>
  select_best(metric = "rmse")

final_lasso <- finalize_workflow(
  wf |> add_model(lasso_tune_spec),
  lowest_rmse
)

saveRDS(final_lasso, here::here("Data", "lasso_final-recipe"))

fit_parsnip <- final_lasso |>
  fit(v5_dat_train) |>
  extract_fit_parsnip()

vis <- fit_parsnip |>
  vi(lambda = lowest_rmse$penalty)

vis |>
  saveRDS(here::here("Data", "lasso_regression.rds"))

vis |>
  filter_out(Importance == 0) |>
  mutate(task = str_split_i(Variable, "_", 1)) |>
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) |>
  ggplot(aes(x = Importance, y = Variable, color = task)) +
  geom_point() +
  labs(y = NULL, color = "Task", title = "Non-Zero Estimates", x = "Absolute Estimate") +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  theme(aspect.ratio = 3)
```

```{r}
lasso_tuned_test_time <- fit_parsnip |> 
  tidy() |>
  filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  inner_join(v5_dat |> select(item, time) |> unique(),
            join_by("term" == "item")) |>
  summarize(.by = task,
            time = sum(time) + get_startup(first(task))) |>
  summarize(time = sum(time)) |>
  pull(time)
```

Time of tuned LASSO test: `r lasso_tuned_test_time |> round(2)` minutes.


```{r}
#| fig-height: 8.5
#| layout-ncol: 2
lasso_estimates |>
  filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  ggplot(aes(y = fct_reorder(term, abs(estimate)), x = abs(estimate),
             color = task)) +
  geom_point() +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  labs(x = "Absolute Estimate", y = NULL, color = "Task",
       title = "No Tuning") +
  theme(aspect.ratio = 3)

vis |>
  filter_out(Importance == 0) |>
  mutate(task = str_split_i(Variable, "_", 1)) |>
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) |>
  ggplot(aes(x = Importance, y = Variable, color = task)) +
  geom_point() +
  labs(y = NULL, color = "Task", title = "Tuning", x = "Absolute Estimate") +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  theme(aspect.ratio = 3)
```



```{r}
cor_including_zeros <- full_join(
  lasso_fit |>
    extract_fit_parsnip() |>
    vi() |>
    select(importance_notuning = Importance, Variable),
  final_lasso |>
    fit(v5_dat_train) |>
    extract_fit_parsnip() |>
    vi(lambda = lowest_rmse$penalty) |>
    select(importance_tuning = Importance, Variable),
  by = "Variable"
) |>
  summarize(cor = cor(importance_notuning, importance_tuning))

cor_not_including_zeros <- full_join(
  lasso_fit |>
    extract_fit_parsnip() |>
    vi() |>
    filter_out(Importance == 0) |>
    select(importance_notuning = Importance, Variable),
  final_lasso |>
    fit(v5_dat_train) |>
    extract_fit_parsnip() |>
    vi(lambda = lowest_rmse$penalty) |>
    filter_out(Importance == 0) |>
    select(importance_tuning = Importance, Variable) ,
  by = "Variable"
) |>
  summarize(cor = cor(importance_notuning, importance_tuning, use = "complete"))
```

Looks like more items included for the tuned model. Relatively similar top items. Correlation between estimates including all the zeros is `r cor_including_zeros |> round(2)` and without all the zeros (dropping those items) is `r cor_not_including_zeros |> round(2)`.


:::

::: {.column width="50%"}
### Elastic Net Tuning

Tuning the `penalty` argument.

```{r}
# tune penalty instead of fixing it
# mixture = 0.5 for elastic net
elasticnet_tune_spec <- linear_reg(penalty = tune(), mixture = 0.5) |>
  set_engine("glmnet")
```

```{r}
#| eval: true
# set.seed(1234)
# 
# # splitting data
# v5_dat_split <- initial_split(v5_dat_wide)
# 
# v5_dat_train <- training(v5_dat_split)
# v5_dat_test <- testing(v5_dat_split)


# v5_dat_boot <- bootstraps(v5_dat_train)

lambda_grid <- grid_regular(penalty(), levels = 50)

doParallel::registerDoParallel()

lasso_grid <- tune_grid(
  wf |> add_model(elasticnet_tune_spec),
  resamples = v5_dat_boot,
  grid = lambda_grid,
  metrics = metric_set(rmse)
)

saveRDS(lasso_grid, here::here("Data", "elastic-net-estimates_tuning.rds"))
```


<!-- ```{r} -->
<!-- #| column: margin -->

<!-- # lasso_grid |> -->
<!-- #   collect_metrics() |> -->
<!-- #   ggplot(aes(penalty, mean, color = .metric)) + -->
<!-- #   geom_errorbar(aes( -->
<!-- #     ymin = mean - std_err, -->
<!-- #     ymax = mean + std_err), -->
<!-- #   alpha = 0.5) + -->
<!-- #   geom_line(linewidth = 1.5, na.rm = T) + -->
<!-- #   facet_wrap(~.metric, scales = "free", nrow = 2) + -->
<!-- #   scale_x_log10() + -->
<!-- #   theme(legend.position = "none") -->
<!-- ``` -->


```{r}
#| fig-height: 10
lasso_grid <- readRDS(here::here("Data", "elastic-net-estimates_tuning.rds"))

lowest_rmse <- lasso_grid |>
  select_best(metric = "rmse")

final_lasso <- finalize_workflow(
  wf |> add_model(elasticnet_tune_spec),
  lowest_rmse
)

saveRDS(final_lasso, here::here("Data", "elastic-net_final-recipe"))

fit_parsnip <- final_lasso |>
  fit(v5_dat_train) |>
  extract_fit_parsnip()

vis <- fit_parsnip |>
  vi(lambda = lowest_rmse$penalty)

vis |>
  saveRDS(here::here("Data", "elastic-net_regression.rds"))

vis |>
  filter_out(Importance == 0) |>
  mutate(task = str_split_i(Variable, "_", 1)) |>
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) |>
  ggplot(aes(x = Importance, y = Variable, color = task)) +
  geom_point() +
  labs(y = NULL, color = "Task", title = "Non-Zero Estimates", x = "Absolute Estimate") +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  theme(aspect.ratio = 3)
```

```{r}
lasso_tuned_test_time <- fit_parsnip |> 
  tidy() |>
  filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  inner_join(v5_dat |> select(item, time) |> unique(),
            join_by("term" == "item")) |>
  summarize(.by = task,
            time = sum(time) + get_startup(first(task))) |>
  summarize(time = sum(time)) |>
  pull(time)
```

Time of tuned elastic net test: `r lasso_tuned_test_time |> round(2)` minutes.

```{r}
#| layout-ncol: 2
#| fig-height: 9.5
lasso_estimates |>
  filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  ggplot(aes(y = fct_reorder(term, abs(estimate)), x = abs(estimate),
             color = task)) +
  geom_point() +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  labs(x = "Absolute Estimate", y = NULL, color = "Task",
       title = "No Tuning") +
  theme(aspect.ratio = 3)

vis |>
  filter_out(Importance == 0) |>
  mutate(task = str_split_i(Variable, "_", 1)) |>
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) |>
  ggplot(aes(x = Importance, y = Variable, color = task)) +
  geom_point() +
  labs(y = NULL, color = "Task", title = "Tuning", x = "Absolute Estimate") +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  theme(aspect.ratio = 3)
```


```{r}
cor_including_zeros <- full_join(
  lasso_fit |>
    extract_fit_parsnip() |>
    vi() |>
    select(importance_notuning = Importance, Variable),
  final_lasso |>
    fit(v5_dat_train) |>
    extract_fit_parsnip() |>
    vi(lambda = lowest_rmse$penalty) |>
    select(importance_tuning = Importance, Variable),
  by = "Variable"
) |>
  summarize(cor = cor(importance_notuning, importance_tuning))

cor_not_including_zeros <- full_join(
  lasso_fit |>
    extract_fit_parsnip() |>
    vi() |>
    filter_out(Importance == 0) |>
    select(importance_notuning = Importance, Variable),
  final_lasso |>
    fit(v5_dat_train) |>
    extract_fit_parsnip() |>
    vi(lambda = lowest_rmse$penalty) |>
    filter_out(Importance == 0) |>
    select(importance_tuning = Importance, Variable) ,
  by = "Variable"
) |>
  summarize(cor = cor(importance_notuning, importance_tuning, use = "complete"))
```

Looks like more items included for the tuned model. Relatively similar top items. Correlation between estimates including all the zeros is `r cor_including_zeros |> round(2)` and without all the zeros (dropping those items) is `r cor_not_including_zeros |> round(2)`.


:::
::::

## LASSO vs Elastic Net

:::: {.columns}
::: {.column width="50%"}
```{r}
#| fig-height: 10
here::here("Data", "lasso_regression.rds") |>
  readRDS() |>
  filter_out(Importance == 0) |>
  mutate(task = str_split_i(Variable, "_", 1)) |>
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) |>
  ggplot(aes(x = Importance, y = Variable, color = task)) +
  geom_point() +
  labs(y = NULL, color = "Task", title = "LASSO", x = "Absolute Estimate") +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  theme(aspect.ratio = 3)
```
:::

::: {.column width="50%"}
```{r}
#| fig-height: 10
here::here("Data", "elastic-net_regression.rds") |>
  readRDS() |>
  filter_out(Importance == 0) |>
  mutate(task = str_split_i(Variable, "_", 1)) |>
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) |>
  ggplot(aes(x = Importance, y = Variable, color = task)) +
  geom_point() +
  labs(y = NULL, color = "Task", title = "Elastic Net", x = "Absolute Estimate") +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  theme(aspect.ratio = 3)
```
:::
::::


## Tuning Elastic Net Mixture

### Elastic Net Tuning

Tuning the `penalty` and `mixture` argument.

```{r}
# tune penalty instead of fixing it
# mixture = 0.5 for elastic net
elasicnet_more_tune_spec <- linear_reg(penalty = tune(), mixture = tune()) |>
  set_engine("glmnet")
```

```{r}
#| eval: true
# set.seed(1234)
# 
# # splitting data
# v5_dat_split <- initial_split(v5_dat_wide)
# 
# v5_dat_train <- training(v5_dat_split)
# v5_dat_test <- testing(v5_dat_split)
# 
# 
# v5_dat_boot <- bootstraps(v5_dat_train)

lambda_grid <- grid_regular(penalty(), levels = 50) |>
  expand_grid(mixture = seq(0.2, 0.8, 0.2))

doParallel::registerDoParallel()

lasso_grid <- tune_grid(
  wf |> add_model(elasicnet_more_tune_spec),
  resamples = v5_dat_boot,
  grid = lambda_grid,
  metrics = metric_set(rmse)
)

saveRDS(lasso_grid, here::here("Data", "elastic-net-estimates_more-tuning.rds"))
```


<!-- ```{r} -->
<!-- #| column: margin -->

<!-- # lasso_grid |> -->
<!-- #   collect_metrics() |> -->
<!-- #   ggplot(aes(penalty, mean, color = .metric)) + -->
<!-- #   geom_errorbar(aes( -->
<!-- #     ymin = mean - std_err, -->
<!-- #     ymax = mean + std_err), -->
<!-- #   alpha = 0.5) + -->
<!-- #   geom_line(linewidth = 1.5, na.rm = T) + -->
<!-- #   facet_wrap(~.metric, scales = "free", nrow = 2) + -->
<!-- #   scale_x_log10() + -->
<!-- #   theme(legend.position = "none") -->
<!-- ``` -->


```{r}
#| fig-height: 10
lasso_grid <- readRDS(here::here("Data", "elastic-net-estimates_more-tuning.rds"))

lowest_rmse <- lasso_grid |>
  select_best(metric = "rmse")

final_lasso <- finalize_workflow(
  wf |> add_model(elasicnet_more_tune_spec),
  lowest_rmse
)

saveRDS(final_lasso, here::here("Data", "elastic-net_more-tuning_final-recipe"))

fit_parsnip <- final_lasso |>
  fit(v5_dat_train) |>
  extract_fit_parsnip()

vis <- fit_parsnip |>
  vi(lambda = lowest_rmse$penalty)

vis |>
  saveRDS(here::here("Data", "elastic-net_regression_more-tuning.rds"))

vis |>
  filter_out(Importance == 0) |>
  mutate(task = str_split_i(Variable, "_", 1)) |>
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) |>
  ggplot(aes(x = Importance, y = Variable, color = task)) +
  geom_point() +
  labs(y = NULL, color = "Task", title = "Non-Zero Estimates", x = "Absolute Estimate") +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  theme(aspect.ratio = 3)
```

```{r}
lasso_tuned_test_time <- fit_parsnip |> 
  tidy() |>
  filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  inner_join(v5_dat |> select(item, time) |> unique(),
            join_by("term" == "item")) |>
  summarize(.by = task,
            time = sum(time) + get_startup(first(task))) |>
  summarize(time = sum(time)) |>
  pull(time)
```

Time of further-tuned elastic net test: `r lasso_tuned_test_time |> round(2)` minutes.


## Test Data Fit Comparison

```{r}
fit_statistics <- c("lasso_final-recipe", "elastic-net_final-recipe", "elastic-net_more-tuning_final-recipe") |>
  map(\(type) here::here("Data", type) |> 
        readRDS() |>
        last_fit(v5_dat_split, metrics = metric_set(rmse, rsq, mae)) |>
        collect_metrics() |>
        mutate(method = type |>
                 str_remove("_final-recipe") |>
                 str_replace_all("_|-", " ") |>
                 str_to_title(),
               .before = everything())) |>
  list_rbind() |>
  select(method, metric = .metric, estimate = .estimate) |>
  pivot_wider(names_from = metric, values_from = estimate)

fit_stat_cols <- fit_statistics |>
  select(-method) |> 
  names()

map(
  fit_stat_cols,
  \(col) list(col = col, idx = which(fit_statistics[[col]] == min(fit_statistics[[col]], na.rm = TRUE)))
) |>
  keep(\(x) length(x$idx) > 0) |>
  reduce(
    \(tbl, x) tab_style(
      tbl,
      style = cell_fill(color = "#e2eee2"),
      locations = cells_body(columns = all_of(x$col), rows = x$idx)
    ),
    .init = gt(fit_statistics)
  ) |>
  fmt_number(decimals = 3) |>
  tab_style(cell_text(weight = "bold"),
            cells_column_labels()) |>
  cols_label_with(columns = -method, str_to_upper) |>
  cols_label_with(columns = method, str_to_title)
```




























