---
title: "LASSO Regression"
date: last-modified
toc: true
format: 
  html:
    code-fold: true
    grid: 
      margin-width: 350px
      gutter-width: 1em
author: Jessica Helmer
knitr:
    opts_chunk: 
      dev: png
      dev.args:
        bg: transparent
---

::: panel-tabset
#### Setup

```{r}
#| message: false
#| code-fold: show
library(tidyverse)
library(tidymodels)
library(vip)

set_theme(theme_classic(base_size = 16, paper = "#eceadf"))
```

#### Reading

```{r}
#| code-fold: show
v5_dat <- readRDS(here::here("Data", "v5_dat.rds"))

startup_costs <- readRDS(here::here("Data", "startup_costs.rds"))
```

#### Functions

```{r}
#| code-fold: show
source(here::here("Scripts", "get_startup.R"))
```

#### Cleaning

```{r}
v5_dat_wide <- v5_dat |>
  # condensing denominator neglect scores down to one score per item pair
  mutate(.by = c(subject_id, item),
         score = ifelse(task == "dn.c" | task == "dn.s",
                        mean(score),
                        score)) |>
  unique() |>
  select(-c(time, task)) |>
  pivot_wider(names_from = "item", values_from = "score",
              id_cols = c(subject_id, sscore)) |>
  # alert alert !!
  drop_na()
```
:::


### LASSO

```{r}
#| message: false
set.seed(12345)

# splitting data
v5_dat_split <- initial_split(v5_dat_wide)

v5_dat_train <- training(v5_dat_split)
v5_dat_test <- testing(v5_dat_split)

# building "recipe" of all items regressed on sscore
v5_dat_recipe <- recipe(sscore ~ ., data = v5_dat_train) |>
  # not including subject_id
  update_role(subject_id, new_role = "ID") |>
  # and z-scoring first
  step_normalize(all_numeric(), -all_outcomes())

# linear regression with glmnet, mixture = 1 to do LASSO
lasso_spec <- linear_reg(penalty = 0.1, mixture = 1) %>%
  set_engine("glmnet")

wf <- workflow() %>%
  add_recipe(v5_dat_recipe)

# fitting model with training data
lasso_fit <- wf %>%
  add_model(lasso_spec) %>%
  fit(data = v5_dat_train)

# extracting estimates
lasso_estimates <- lasso_fit %>%
  extract_fit_parsnip() %>%
  tidy()

# display estimates
# lasso_estimates |> 
#   filter_out(when_any(estimate == 0, 
#                       term == "(Intercept)")) |>
#   mutate(task = term |> str_split_i("_", 1)) |>
#   select(task, item = term, estimate)
```

```{r}
lasso_estimates |>
  filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  ggplot(aes(y = fct_reorder(term, abs(estimate)), x = abs(estimate),
             color = task)) +
  geom_point() +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  labs(x = "Absolute Estimate", y = NULL, color = "Task",
       title = "Non-Zero Estimates") +
  theme(aspect.ratio = 3)
```

```{r}
lasso_test_time <- lasso_estimates |> filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  inner_join(v5_dat |> select(item, time) |> unique(),
            join_by("term" == "item")) |>
  summarize(.by = task,
            time = sum(time) + get_startup(first(task))) |>
  summarize(time = sum(time)) |>
  pull(time)
```

Time of LASSO test: `r lasso_test_time |> round(2)` minutes.


### LASSO Tuning

Trying some tuning with the `penalty` argument to see if it makes a difference!

```{r}
# tune penalty instead of fixing it
tune_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")
```

```{r}
#| eval: true
set.seed(1234)

# splitting data
v5_dat_split <- initial_split(v5_dat_wide)

v5_dat_train <- training(v5_dat_split)
v5_dat_test <- testing(v5_dat_split)


v5_dat_boot <- bootstraps(v5_dat_train)

lambda_grid <- grid_regular(penalty(), levels = 50)

doParallel::registerDoParallel()

lasso_grid <- tune_grid(
  wf %>% add_model(tune_spec),
  resamples = v5_dat_boot,
  grid = lambda_grid,
  metrics = metric_set(rmse)
)

saveRDS(lasso_grid, here::here("Data", "lasso-estimates_tuning.rds"))
```

```{r}
#| column: margin
lasso_grid <- readRDS(here::here("Data", "lasso-estimates_tuning.rds"))

lasso_grid %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_errorbar(aes(
    ymin = mean - std_err,
    ymax = mean + std_err),
  alpha = 0.5) +
  geom_line(linewidth = 1.5, na.rm = T) +
  facet_wrap(~.metric, scales = "free", nrow = 2) +
  scale_x_log10() +
  theme(legend.position = "none")
```

```{r}
lowest_rmse <- lasso_grid %>%
  select_best(metric = "rmse")

final_lasso <- finalize_workflow(
  wf %>% add_model(tune_spec),
  lowest_rmse
)

fit_parsnip <- final_lasso %>%
  fit(v5_dat_train) %>%
  extract_fit_parsnip()

vis <- fit_parsnip %>%
  vi(lambda = lowest_rmse$penalty)

vis |>
  saveRDS(here::here("Data", "lasso_regression.rds"))

vis %>%
  filter_out(Importance == 0) |>
  mutate(task = str_split_i(Variable, "_", 1)) |>
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) %>%
  ggplot(aes(x = Importance, y = Variable, color = task)) +
  geom_point() +
  labs(y = NULL, color = "Task", title = "Non-Zero Estimates", x = "Absolute Estimate") +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  theme(aspect.ratio = 3)
```

```{r}
lasso_tuned_test_time <- fit_parsnip |> 
  tidy() |>
  filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  inner_join(v5_dat |> select(item, time) |> unique(),
            join_by("term" == "item")) |>
  summarize(.by = task,
            time = sum(time) + get_startup(first(task))) |>
  summarize(time = sum(time)) |>
  pull(time)
```

Time of tuned LASSO test: `r lasso_tuned_test_time |> round(2)` minutes.


```{r}
#| layout-ncol: 2
#| fig-height: 6
lasso_estimates |>
  filter(estimate != 0 & term != "(Intercept)") |>
  mutate(task = str_split_i(term, "_", 1)) |>
  ggplot(aes(y = fct_reorder(term, abs(estimate)), x = abs(estimate),
             color = task)) +
  geom_point() +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  labs(x = "Absolute Estimate", y = NULL, color = "Task",
       title = "No Tuning") +
  theme(aspect.ratio = 3)

vis %>%
  filter_out(Importance == 0) |>
  mutate(task = str_split_i(Variable, "_", 1)) |>
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) %>%
  ggplot(aes(x = Importance, y = Variable, color = task)) +
  geom_point() +
  labs(y = NULL, color = "Task", title = "Tuning", x = "Absolute Estimate") +
  annotate("segment", x = 0.05, y = -Inf, yend = Inf, alpha = 0.2) +
  scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
  coord_cartesian(xlim = c(0, 0.1)) +
  theme(aspect.ratio = 3)
```

```{r}
cor_including_zeros <- full_join(
  lasso_fit %>%
    extract_fit_parsnip() |>
    vi() %>%
    select(importance_notuning = Importance, Variable),
  final_lasso %>%
    fit(v5_dat_train) %>%
    extract_fit_parsnip() %>%
    vi(lambda = lowest_rmse$penalty) %>%
    select(importance_tuning = Importance, Variable),
  by = "Variable"
) |>
  summarize(cor = cor(importance_notuning, importance_tuning))

cor_not_including_zeros <- full_join(
  lasso_fit %>%
    extract_fit_parsnip() |>
    vi() %>%
    filter_out(Importance == 0) |>
    select(importance_notuning = Importance, Variable),
  final_lasso %>%
    fit(v5_dat_train) %>%
    extract_fit_parsnip() %>%
    vi(lambda = lowest_rmse$penalty) %>%
    filter_out(Importance == 0) |>
    select(importance_tuning = Importance, Variable) ,
  by = "Variable"
) |>
  summarize(cor = cor(importance_notuning, importance_tuning, use = "complete"))
```

Looks like more items included for the tuned model. Relatively similar top items. Correlation between estimates including all the zeros is `r cor_including_zeros` and without all the zeros (dropping those items) is `r cor_not_including_zeros`.

