---
title: "{xgboost}"
subtitle: "With {xgboost}"
date: last-modified
toc: true
format: 
  html:
    code-fold: true
    grid: 
      margin-width: 350px
      gutter-width: 1em
author: Jessica Helmer
knitr:
    opts_chunk: 
      dev: png
      dev.args:
        bg: transparent
---

::: panel-tabset
#### Setup

```{r}
#| message: false
#| code-fold: show
library(tidyverse)
library(tidymodels)
library(finetune)
library(vip)
```

#### Reading

```{r}
#| code-fold: show
v5_dat <- readRDS(here::here("Data", "v5_dat.rds"))

startup_costs <- readRDS(here::here("Data", "startup_costs.rds"))
```

#### Functions

```{r}
#| code-fold: show
source(here::here("Scripts", "get_startup.R"))
```

#### Cleaning
```{r}
v5_dat_wide <- v5_dat |>
  # condensing denominator neglect scores down to one score per item pair
  mutate(.by = c(subject_id, item),
         score = ifelse(task == "dn.c" | task == "dn.s",
                        mean(score),
                        score)) |>
  unique() |>
  select(-c(time, task)) |>
  pivot_wider(names_from = "item", values_from = "score",
              id_cols = c(subject_id, sscore)) |>
  # alert alert !!
  drop_na()
```

:::


## Trying `{xgboost}`

Splitting data into train and test.

```{r}
v5_dat_split <- v5_dat_wide |>
  select(-subject_id) |>
  initial_split()

v5_dat_train <- training(v5_dat_split)
v5_dat_test <- testing(v5_dat_split)

v5_dat_folds <- vfold_cv(v5_dat_train, v = 5)
```

Specifying model (S-score predicted by everything else) and normalizing all predictors.

```{r}
v5_dat_recipe <- recipe(sscore ~ ., data = v5_dat_train) |>
  step_normalize(all_numeric(), -all_outcomes())
```

Specifying hyperparameters and which are going to be tuned.

```{r}
xgb_spec <- boost_tree(
    trees = 1000,
    tree_depth = tune(),
    min_n = tune(),
    mtry = tune(),
    sample_size = tune(),
    learn_rate = tune()
  ) %>%
  set_engine("xgboost") |>
  set_mode("regression")

xgb_wf <- workflow(v5_dat_recipe, xgb_spec)
```

### Hyperparameter Tuning

Defining grid of candidate hyperparameter combinations.

```{r}
xgb_grid <- grid_space_filling(
    tree_depth(c(5L, 10L)),
    min_n(c(1L, 40L)),
    mtry(c(1L, 90L)),
    sample_prop(c(0.5, 1.0)),
    learn_rate(c(-3, -1)),
    size = 100)

xgb_grid
```

Five-fold cross-validation for tuning.

```{r}
#| eval: false
doParallel::registerDoParallel()

set.seed(234)
xgb_rs <- tune_race_anova(
    xgb_wf,
    v5_dat_folds,
    grid = xgb_grid,
    control = control_race(verbose_elim = TRUE))

if (!file.exists(here::here("Data", "xgboost"))) dir.create(here::here("Data", "xgboost"))
saveRDS(xgb_rs, here::here("Data", "xgboost", "xgb_rs.rds"))
```

```{r}
#| column: margin
xgb_rs <- readRDS(here::here("Data", "xgboost", "xgb_rs.rds"))

plot_race(xgb_rs) +
  theme_classic(base_size = 20)
```
Highest-performing hyperparameter combinations

```{r}
show_best(xgb_rs)
```

### Fitting Final Model

```{r}
xgb_last <-
  xgb_wf %>%
  finalize_workflow(select_best(xgb_rs, metric = "rmse")) %>%
  last_fit(v5_dat_split)

xgb_last
```

Importance of items

```{r}
#| fig-column: margin
#| fig-height: 16

extract_workflow(xgb_last) %>%
  extract_fit_parsnip() |>
  vi() |>
  saveRDS(here::here("Data", "xgboost", "xgb_item-importance.rds"))


extract_workflow(xgb_last) %>%
  extract_fit_parsnip() %>%
  vip(geom = "point", num_features = 100) +
  theme_classic(base_size = 16) +
  theme(aspect.ratio = 4,
        panel.grid.major.x = element_line(linewidth = 0.2))
```












