---
title: "2PL"
---

```{r}
#| message: false
library(tidyverse)
```


## 2PL Modeling Strategy

Included: Denominator Neglect, Decision Rules

$$ logodds(P(Y = 1)) = a (\theta - b) $$, where $b$ represents the item's *difficulty* and $a$ represents the item's *discrimination*.

*Difficulty* corresponds to the location of the item response curve, indicating how... difficult... the item is to get correct. Higher difficulty corresponds to harder items, and difficulty values scale with $\theta$: an item with a difficultly value of $b = 0$ means that an examinee with average skill $\theta = 0$ would have a 50% chance of a correct answer. 

*Discrimination* corresponds to the steepness of the item response curve and indicates how well items... discriminate... among $\theta$ levels. Higher values indicate steeper slopes which better discriminate among skill levels.

```{r}
expand_grid(a = c(.5, 1, 1.5),
            b = c(-2, 0, 2),
            theta = seq(-5, 5, length.out = 100)) |>
  mutate(p = plogis(a * (theta - b))) |>
  ggplot(aes(x = theta, y = p)) +
  geom_line() +
  scale_x_continuous(breaks = c(-3, 0, 3)) +
  scale_y_continuous(breaks = c(0, .5, 1)) +
  guides(x = guide_axis(cap = "both"),
         y = guide_axis(cap = "both")) +
  facet_grid(cols = vars(a), rows = vars(b),
             labeller = label_both,
            axes = "all") +
  theme_classic() +
  theme(strip.background = element_blank())
```
#### 2PL Stan Code

```{r setup}
#| include: false
library(tidyverse)

knitr::read_chunk(
  lines = readLines(here::here("Models", "2pl-code.stan")),
  label = "stan_code")
```

### Parameter Estimates to Curves

To get item response and information curves from the Stan model's parameter estimates, we first converted the item paramter estimates to probabilites of a correct answer for a range of theoretical $\theta$ values.

$$ P(Y = 1 | \theta) \bigg|_{\theta \in [-5, 5]} = \frac{e^{\hat{a} (\theta - \hat{b})}}{1 + e^{\hat{a} (\theta - \hat{b})}} $$

```{r}
dat <- expand_grid(a = c(.5, 1, 1.5),
            b = c(-2, 0, 2),
            theta = seq(-5, 5, length.out = 100)) |>
  mutate(p = plogis(a * (theta - b)))
```


We plotted those predicted probabilities of a correct answer for the theoretical $\theta$ values.

```{r}
dat |>
  ggplot(aes(x = theta, y = p)) +
  geom_line() +
  scale_x_continuous(breaks = c(-3, 0, 3)) +
  scale_y_continuous(breaks = c(0, .5, 1)) +
  guides(x = guide_axis(cap = "both"),
         y = guide_axis(cap = "both")) +
  facet_grid(cols = vars(a), rows = vars(b),
             labeller = label_both,
            axes = "all") +
  theme_classic() +
  theme(strip.background = element_blank())
```

To get information curves for an item, we multiply the probability of a correct answer by the probability of an incorrect answer and by the square of the discrimination parameter.

$$ I(\theta) \bigg|_{\theta \in [-5, 5]} = a^2 \times P(Y = 1) \times P(Y = 0) = a^2 \times P(Y = 1) \times (1 - P(Y = 1)) $$

And plot the information function values over theoretical values of $\theta$.

```{r}
dat |>
  mutate(I = a^2 * p * (1 - p)) |>
  ggplot(aes(x = theta, y = I)) +
  geom_line() +
  scale_x_continuous(breaks = c(-3, 0, 3)) +
  scale_y_continuous(breaks = c(0, .25)) +
  guides(x = guide_axis(cap = "both"),
         y = guide_axis(cap = "both")) +
  facet_grid(cols = vars(a), rows = vars(b),
             labeller = label_both,
            axes = "all") +
  theme_classic() +
  theme(strip.background = element_blank())
```
